---
title: "Assignment week7"
author: "Eunkyu Hahm"
date: "10/11/2019"
output:
  html_document:
    df_print: paged
    highlight: tango
    theme: united
    pandoc_args: [
      "--title-prefix", "Foo",
      "--id-prefix", "Bar"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# *Loading packages*  
Firstly, let's load necessary packages. XML package is used to parse XML and html file, and jsonlite is used to parse json file. 
```{r}
library(RCurl)
library(XML)
library(jsonlite)
library(DT)
library(stringr)
library(tidyr)
library(dplyr)
```

## HTML

Let's read html file from github with using getURL() function and then read html table. Since the clasee of html_book is a list, I used data.frame() function in order to convert to data frame class.
```{r}
html <- getURL("https://raw.githubusercontent.com/ekhahm/datascience/master/week7/books.html")
html_book<- readHTMLTable(html)
html_book
class(html_book)
html_df <- data.frame(html_book)
html_df
```
Now, we are taking a look of the html data table. 
```{r}
options(DT.options = list(dom = 't', scrollX = TRUE))
datatable(html_df)
```

## XML  
Let's get xml file from github with using getURL() function and then parse xml table with xmlParse(). I used getNodeSet() function to find matching each node in an xml treen and then change the class to dataframe. Lastly set the names in an object for each node. 
```{r}
xml <-getURL("https://raw.githubusercontent.com/ekhahm/datascience/master/week7/books.xml")
xml_book <- xmlParse(xml)

a <- setNames(xmlToDataFrame(node = getNodeSet(xml_book, "//root/book/Title")), "Title")
b <- setNames(xmlToDataFrame(node = getNodeSet(xml_book, "//root/book/Author")), "Author")
c <- setNames(xmlToDataFrame(node = getNodeSet(xml_book, "//root/book/ISBN")), "ISBN")
d <- setNames(xmlToDataFrame(node = getNodeSet(xml_book, "//root/book/Genre")), "Genre")

xml_book <- cbind(a,b,c,d)
xml_book
```

Now, we are taking a look of xml data table. 
```{r}
datatable(xml_book)
```


## JSON

Let's read JSON file from github with using getURL() function and then convert r object from Json using fromJSON . Since the clasee of json_book is a list, I used data.frame() function in order to convert to data frame class.
```{r}
json <- getURL("https://raw.githubusercontent.com/ekhahm/datascience/master/week7/books.json")
json_book <- fromJSON(json)
json_book
json_df <- data.frame(json_book)
json_df
```
Now we are looking at the data table of json_df
```{r}
datatable(json_df)
```

# *Conclusion* 

The three data frames from html, xml, and json files look identical. 