---
title: 'Assignmnet 3 '
author: "Eunkyu Hahm"
date: "9/12/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to problems 3 and 4 from chapter 8 of Automated Data Collection in R.  Problem 9 is extra credit.  You may work in a small group, but please submit separately with names of all group participants in your submission.

Here is the referenced code for the introductory example in #3:

```{r}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

raw.data
```


##Loading stringr package
```{r stringr}
library(stringr)
```

## Problem #3 
Copy the introductory example. The vector name stores the extracted names. 

```{r}
names <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
names
```

(a) Use the tools of this chapter to rearrange the cector so that all elements conform to the standard first_name last_name. 

Removing titles and middle name:
```{r title}
no_title <- sub("[A-z]{1, }\\.","",names )
no_title
```

First name and Last name
```{r first_last }
first_last <- sub("(\\w+),\\s+(\\w+)","\\2 \\1", no_title )
first_last
```
(b) Construct a logical vector indicating whether a character has a title(i.e.Rev, and Dr. )
```{r title vector}
title_name <- str_detect(names, "[A-z]{2,}[.]")
title_name
```

(c) Costruct a logical vector indicating whether a character has a second name. 
```{r second_name}
#remove title 
non_title <- sub("[A-z]{2, }\\.","",names )
second_name <- str_detect(non_title, "[A-z]{1,}\\.")
second_name
```

## Problem #4
Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression. 

(a) [0-9]+\\$
 one digit or more digits followed by a dollar sign. 
```{r (a)}
a <- c("4300$", "120$", "$10", "$1")
str_count(a, "[0-9]+\\$")
```
(b) \\b[a-z]{1,4}\\b

small letter a to z at lease 1 time but not more than 4 times with word edge each side. 
```{r b}
b <- c("1abc", "terminal", "size", "p", "1001")
str_detect(b, "\\b[a-z]{1,4}\\b")
```

(c) .*?\\.txt$
start with any character followed by end with ".txt" 

```{r c}
c <- c(".txt", "8.txt", "case.txt", "txt", "123")
str_extract(c, ".*?\\.txt$")
```


(d) \\d{2}/\\d{2}/\\d{4}
two digits followed by forward slash("/") and two digits followed by forward slash("/") and followed by any four digits 

```{r d}
d <- c("10/30/109", "11/31/2019", "3-19-2001", "1985/03/14", "4/5/1000")
str_detect(d, "\\d{2}/\\d{2}/\\d{4}")
```


(e) <(.+?)>.+?</\\1>
any items in between < > followed by any items and backreferecning the first items in < > with forward slash("/")
```{r e}
e <- c("<random>Description</random>", "<random>Description<random>", "<tools>Useage</tools>", "<tools>Useage</tool>", "size", "<p>123</p>", "<1>False</1>")
str_extract_all(e, "<(.+?)>.+?</\\1>")
```

