---
title: "Assignment 2 - SQL and R"
author: "Eunkyu Hahm"
date: "9/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Choose six recent popular movies. Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5. Take the results (observations) and store them in a SQL database. Load the information into an R dataframe.
Your deliverables should include your SQL scripts and your R Markdown code, posted to GitHub.


## Loading RMySQL
```{r Loading}
library(RMySQL)
```

## Connecting mysql server
```{r}
drv <- dbDriver("MySQL")
con <- dbConnect(drv, user= "root", password = "gkadmsrb0", dbname = "movies", host="localhost")
```

## Take the observations and store them in a sql databse

1. Movie table 

```{r}
movie <- "SELECT * from movies;"
movie_title <- dbGetQuery(con, movie)
movie_title
```

2. Participants table 
```{r}
subjects <- "SELECT * from subjects;"
participants <- dbGetQuery(con, subjects)
participants
```
3. Rating table

```{r}
ratings <- "SELECT * from ratings;"
movies_ratings <- dbGetQuery(con, ratings)
movies_ratings
```

5. Summary of movie ratings
Average ratings of each movie. The movie Avengers:endgame has the highest rating. 
```{r}
summary  <- "select title, avg(ratings) from movies, ratings
where movies.mID = ratings.mID
group by movies.title
order by avg(ratings);"
sum_rating <- dbGetQuery(con, summary)
sum_rating
```