---
title: "Week5 assignment"
author: "Eunkyu Hahm"
date: "9/27/2019"
output:
  html_document:
    df_print: paged
    highlight: kate
    number_sections: yes
    theme: united
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment – Tidying and Transforming Data

|        |         | Los Angeles | Phoenix | San Diego | San Francisco | Seattle |
|--------|---------|-------------|---------|-----------|---------------|---------
|ALASKA  | on time | 497         |221      |212        |503            |1841     |
|        | delayed |62           |12       |20         |102            |305      |
|AM WEST | on time |694          |4840     |383        |320            |201      |
|        | delayed |117          |415      |65         |129            |61       |

Table 1: The chart above describes arrival delays for two airlines across five destinations.


## Loading packeges 
Firstly, we wil load packages that are used for data tidying and transforming  

* Two packages  
  + tidyr
  + dplyr 
  + ggplot2
  
```{r loading packages}
library(tidyr)
library(dplyr)
library(ggplot2)
```

## Create a .CSV file that incldue the table 1 information 

In order to write .CSV file, we begin with builing a table and put the values in it. And then write out .CSV file. 
```{r csv}
#create the table in r 
table1 <- rbind(c(NA, NA, "Los_Angeles", "Phoenix", "San Diego", "San Francisco", "Seattle"),
          c("ALASKA","on time",497,221,212,503,1841),
          c(NA, "delayed", 62,12,20,102,305),
          c("AM WEST", "on time", 694, 4840,383,320,201),
          c(NA, "delayed", 117,415,65,129,61))

# write out .csv file
write.csv(table1, file = "Flight_arrival_dealy.csv")

```

## Read the information from the .csv file above into R

Let's read the .CSV file that we create above. 
```{r read csv}
delay <- read.csv("Flight_arrival_dealy.csv",header = TRUE, stringsAsFactors = FALSE)
delay
```

## Modify the table 

The table from .CSV file is a little untidy and messy. There are no headers and two NA values that has to be replaced. It will be modified with adding column names, removing unnecessary row and column, and filling NAs with values
```{r}
# check current column names 
colnames(delay) = delay[1,]

# remove the first column and row
delay1= delay[-1, ]
delay2= delay1[ ,-1]

# change column names 
colnames(delay2)[1] <- "Airline"
colnames(delay2)[2] <- "Arrival_Status"

# fill NA with Airline name on "delayed" row 
delay2[2,1] <- "ALASKA"
delay2[4,1] <- "AM WEST"
delay2
```

## Perform analysis to compare the arrival delays for the two airlines with using *tidyr* and *dplyr*.  

Let's take a look at the table above and find the variables.   

* Variables in the table  
  + Airline  
  + Arrival status  
  + City  
  + Count 

  
### Collapse multiple columns into four columns; (Airline, Arrival_Status, City, Count)

Since the format of delay2 table is untidy, let's re-structure the table into tidy format with using gather() function that collapse multibple comlumns into four columns. 

```{r collapse columns}
delay_new <- delay2 %>%
              gather(city, count, c("Los_Angeles":"Seattle"))
delay_new

```

### Return rows with matching specific airline and order by arrival status 

Let's divide into two table according to Airline; Alaska and Am west. To extract the rows of interest, filter() function is used and then arrange() function is used to make an order by arrival status. We will use the pipe operate here because we will pipe the results after one expression into a second expression.

```{r matching conditions}
## return all rows with matching ALASKA airline
alaska <- delay_new %>%
            filter(Airline == "ALASKA") %>%
            arrange(Arrival_Status)
## return all rows with matching AM WEST airline
amwest <- delay_new %>%
            filter(Airline == "AM WEST") %>%
            arrange(Arrival_Status)

alaska
amwest
```  


### Graph expression   

Now we are creating barplot by using ggplot2 package in order to compare observations of each city. Because the total of flights are different between citites, we will generate stacked barplot with normalized height, so that it will be easier to compare the data. 

```{r barplot}
# take each table's information
str(alaska)
str(amwest)

# change count valiable from charactor to numeric
ggplot(alaska, aes(fill= Arrival_Status, x = city, y = as.numeric(count)), ylim(0,5000)) +
  # stacked barplot with normalized height
  geom_bar(stat = "identity", position = "fill")+
  # change title of barplot
  ggtitle("ALASKA Airline") +
  # change x-axis name
  xlab("city") +
  # change y-axis name
  ylab("count")

# change count valiable from charactor to numeric  
ggplot(amwest, aes(fill= Arrival_Status, x = city, y = as.numeric(count)), ylim(0,5000)) +
  # stacked barplot with normalized height
  geom_bar(stat = "identity", position = "fill")+
  # change title of barplot
  ggtitle("AM WEST Airline") +
  # change x-axis name
  xlab("city") +
  # change y-axis name
  ylab("count")

```

### Calculating percentage of delayed flights for each airline   

We will calculate how much proportions of delayed flights and on time flights for each airline. 
Each airline table is grouped by arrival_status, calculate total count for each arrival status , and then add a column for percentage of each case. 

```{r}
#Alaska airline
# make count variable type into numeric 
alaska$count <- as.numeric(alaska$count)
alaska1 <- alaska %>%
            group_by(Arrival_Status) %>%
            summarise(total = sum(count)) %>%
            mutate(sum(total)) %>%
            mutate(percent = round(total/sum(total)*100))
alaska1

#Amwest airline
amwest$count <- as.numeric(amwest$count)
amwest1 <- amwest %>%
            group_by(Arrival_Status) %>%
            summarise(total = sum(count)) %>%
            mutate(sum(total)) %>%
            mutate(percent = round(total/sum(total)*100))
amwest1

```


## Conclusion   

1. According to the ALASKA and AM WEST barplots, San Francisco airport has the highest delayed rates among five cities for both airlines.

2. According to the ALASKA and AM WEST barplots, Phoenix airpot has the lowest delayed rates among five cities for both airlines. 

3. Based on the alaska1 and amwest1 tables, Alaska airline has higher delayed rate in the total number of flights than AM WEST airline. The delayed rate for Alaska is 13% and the rate for AM WEST is 11%. 
But the total number of flights are different which Alaska has total 3775 flights and AM WEST has total 7225 flights. 
