---
title: "Project 1"
author: "Eunkyu Hahm"
date: "9/14/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project 1

In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605
1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

If you have questions about the meaning of the data or the results, please post them on the discussion forum. Data science, like chess, is a game of back and forth...
The chess rating system (invented by a Minnesota statistician named Arpad Elo) has been used in many other contexts, including assessing relative strength of employment candidates by human resource departments.

You may substitute another text file (or set of text files, or data scraped from web pages) of similar or greater complexity, and create your own assignment and solution. You may work in a small team. All of your code should be in an R markdown file (and published to rpubs.com); with your data accessible for the person running the script.

## Understand the project
1. Read given text file.
2. create an R Markdaown file that generate .csv file.
3. The .csv file includes  
   Player's name  
   Player's state  
   Total number of points  
   Player's pre-rating  
   Player's average pre chess rating of opponents  
   (= total pre-tourament opponents' ratings divided by the total number of games played)  
  

## Read text file
```{r read text}
tournament <- read.table("/Users/ekhahm/desktop/tournamentinfo.txt",sep="\n")
tournamentinfo <- tournament[-c(1:4),]
tournamentinfo
```

## Load package
```{r stringr}
library(stringr)
```

## Extract information from text file

```{r player information}
#Extract players' names
playername<-unlist(str_extract_all(tournamentinfo, "[A-Z]+\\s[A-Z]+\\s?[A-Z]*\\s[A-Z]*"))

#Extract players' states
playerstate <- unlist(str_extract_all(tournamentinfo, "^\\s+[A-Z]{2}"))
playerstates <- str_remove_all(playerstate, "\\s+")

#Extract total number of points
totalpoint <- unlist(str_extract_all(tournamentinfo, "\\|[[:digit:]]+\\.[[:digit:]]"))
totalpoints <- str_remove_all(totalpoint, "\\|")

#Extract players' pre-ratings
pre_rating <- unlist(str_extract_all(tournamentinfo, "R:+\\s+[[:digit:]]{1,}"))
pre_ratings <- str_remove_all(pre_rating, "R:+\\s+")

#Combine them into data.frame
tournament_table <- data.frame(number=seq.int(1,64),playername, playerstates, totalpoints, pre_ratings)
tournament_table
```


## Calculate players' averages pre chess rating of opponents


```{r result_match_table}
# Convert text file into dataframe
tournamentinfo1 <- as.data.frame(matrix(unlist(str_split(tournamentinfo, "\\|")), nrow= 64, byrow=T), stringAsFactors= FALSE)
 
# Select columns of interest
result_match <- data.frame(tournamentinfo1$V1, tournamentinfo1$V4, tournamentinfo1$V5, tournamentinfo1$V6,tournamentinfo1$V7, tournamentinfo1$V8, tournamentinfo1$V9,tournamentinfo1$V10)

# Change column names
names(result_match) <- c("number", "R1", "R2", "R3", "R4", "R5", "R6", "R7")

# Extract only numbers
result_match$R1 <- str_extract(result_match$R1, "[:digit:]+")
result_match$R2 <- str_extract(result_match$R2, "[:digit:]+")
result_match$R3 <- str_extract(result_match$R3, "[:digit:]+")
result_match$R4 <- str_extract(result_match$R4, "[:digit:]+")
result_match$R5 <- str_extract(result_match$R5, "[:digit:]+")
result_match$R6 <- str_extract(result_match$R6, "[:digit:]+")
result_match$R7 <- str_extract(result_match$R7, "[:digit:]+")
result_match
```

```{r}
# Calculate average pre chess rating of opponents
avg_pre_rating <- array(0, dim=nrow(tournament_table))
for ( i in 1:64){
      avg_pre_rating[i]<- floor(mean(c(as.numeric(str_trim(tournament_table$pre_ratings[as.numeric(result_match$R1[i])])),
            as.numeric(str_trim(tournament_table$pre_ratings[as.numeric(result_match$R2[i])])),
            as.numeric(str_trim(tournament_table$pre_ratings[as.numeric(result_match$R3[i])])),
            as.numeric(str_trim(tournament_table$pre_ratings[as.numeric(result_match$R4[i])])),
            as.numeric(str_trim(tournament_table$pre_ratings[as.numeric(result_match$R5[i])])),
            as.numeric(str_trim(tournament_table$pre_ratings[as.numeric(result_match$R6[i])])),
            as.numeric(str_trim(tournament_table$pre_ratings[as.numeric(result_match$R7[i])]))), na.rm = TRUE))
};       
# Add average_pre_rating column on tournament_table 
tournament_table$avg_pre_rating <- avg_pre_rating
tournament_table
```

## Generate the r file into .csv file
```{r csv file}
write.csv(tournament_table, file = "Data607project1.csv")
```
