---
title: "Data605 Final project"
author: "Eunkyu Hahm"
date: "5/17/2020"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: True
      smooth_scroll: True
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(matrixcalc)
library(MASS)
library(Rmisc)
```

## Problem 1. 
Using R, generate a random variable X that has 10,000 random uniform numbers from 1 to N, where N can be any number of your choosing greater than or equal to 6.  Then generate a random variable Y that has 10,000 random normal numbers with a mean of μ=σ=(N+1)/2.  

```{r}
set.seed(123)
N <- 6
n <- 10000
mu <- sigma <- (N + 1)/2
df <- data.frame(X = runif(n, min = 1, max = N), Y = rnorm(n, mean = mu, sd = sigma))
summary(df$X)
summary(df$Y)
```

### Probability.   
Calculate as a minimum the below probabilities a through c.  Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable.  Interpret the meaning of all probabilities.

```{r}
x <- median(df$X)
y <- quantile(df$Y, 0.25)
```
a.   P(X>x | X>y)		
```{r}
total <- nrow(df)
a <- nrow(subset(df, X>y))/total
a1 <- nrow(subset(df, X>x & Y>y))/total
a1/a
```

b.  P(X>x, Y>y)
```{r}
b <- nrow(subset(df, X>x& Y>y))/total
b
```
c.  P(X<x | X>y)	
```{r}
c <- nrow(subset(df, X<x& X>y))/total
c
```
### P(X>x and Y>y)=P(X>x)P(Y>y)
Investigate whether P(X>x and Y>y)=P(X>x)P(Y>y) by building a table and evaluating the marginal and joint probabilities.  
```{r}
pxgx <- nrow(subset(df, X>x))/total
pygy <- nrow(subset(df, Y>y))/total
pxgxygy<- nrow(subset(df, X>x & Y>y))/total

pxpy <- pxgx *pygy

round(pxgxygy, 2)
round(pxpy, 2)
```
Answer: P(X>x and Y>y)=P(X>x)P(Y>y) is true

### Table 
```{r}
table <- cbind(pxgx, pygy, pxpy, pxpy)
colnames(table) <- c( "P(X>x)", "P(Y>y)", "P(X>x)P(Y>y)", "P(X>x & Y>y)")
table
```

```{r}
a1 <- subset(df, X > x)
b1 <- subset(df, Y > y)
le_x <- subset(df, X <= x)
le_y <- subset(df, Y <= y)
table <- matrix(c(nrow(a1), nrow(b1), nrow(le_x), nrow(le_y)), 2, 2,
                dimnames = list(c("x", "y"),
                                c("X > x, Y > y", "X <= x, Y <= y")))
table
```


### Fisher's exact/Chi square Test
H0 : X and Y are independent
HA : Y and Y are not independent 
Fisher's exact test
```{r}
fisher.test(table)
```
Chi squre test
```{r}
chisq.test(table)
```
P-value is less than 0.05. The null hypothesis is rejected. Therefore, X and Y are not indepedent.

## Problem 2. 
You are to register for Kaggle.com (free) and compete in the House Prices: Advanced Regression Techniques competition.  https://www.kaggle.com/c/house-prices-advanced-regression-techniques .  I want you to do the following.

### Descriptive and Inferential Statistics  
```{r}
train <- read.csv("https://raw.githubusercontent.com/ekhahm/datascience/master/Data605/train.csv")
head(train)

```
#### 1. descriptive statistics
Provide univariate descriptive statistics and appropriate plots for the training data set.

Variable 1:Overall Quality
```{r}
summary(train$OverallQual)
```

```{r}
par(mfrow=c(1,2))
boxplot(train$OverallQual, main="Boxplot of Overall Quality")
hist(train$OverallQual, main= "Histogram of Overall Quality", col= "blue")
```
Independent variable 2 : Year built
```{r}
summary(train$YearBuilt)
```
```{r}
par(mfrow=c(1,2))
boxplot(train$YearBuilt, main="Boxplot of Year built")
hist(train$YearBuilt, main= "Histogram of Year Built", col= "red")
```


Dependent variable : Sale price
```{r}
summary(train$SalePrice)
```

```{r}
par(mfrow=c(1,2))
boxplot(train$SalePrice, main="Boxplot of sale price")
hist(train$SalePrice, main= "Histogram of sale price", col= "red")
```


####2. Scatterplot  

```{r}
plot(train$OverallQual, train$SalePrice, xlab= "Overall quality", ylab="Sale price", main= "scatterplot of Overall Quality vs. Sale Price")
```

```{r}
plot(train$YearBuilt, train$SalePrice, xlab= "Year Built", ylab="Sale price", main= "scatterplot of Year Built vs. Sale Price")
```

#### 3. Corrleation matrix
```{r}
corDF <- train[c("SalePrice", "OverallQual", "YearBuilt")]
corMatrix <- cor(corDF, use = "complete.obs")
print(corMatrix)
```
#### 4. Test hypothesis
```{r}
cor.test(train$OverallQual, train$SalePrice, conf.level = 0.8)
```
The correlation value is 0.79. It indicates there is a correlation between overall quality and sale price. P-value is less than 0.05. It can be concluded that two varialbes are correlated. 

```{r}
cor.test(train$YearBuilt, train$SalePrice, conf.level = 0.8)
```
The correlation value is 0.522. It indeicates there is moderate correlation between Year built and sale price.

### Linear algebra and Correlation
Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.) Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix. Conduct LU decomposition on the matrix.  

Precision matrix
```{r}
p_matrix <- solve(corMatrix)
p_matrix
```
Multiply the correlation matrix by the precision matrix
```{r}
multiply <- round(corMatrix %*% p_matrix)
multiply
```

Multiply the precision matrix by the correlation matrix
```{r}
multiply1 <- round(p_matrix %*% corMatrix)
multiply1
```
LU decomposition on the matrix.  
```{r}
decomposition <- lu.decomposition(corMatrix)
decomposition
corMatrix
```
### Calculus Based Probaility & Statistics
```{r}
X <- train$OverallQual[!is.na(train$OverallQual)]
fd <- fitdistr(X, "gamma")
hist(X, breaks=40, prob=TRUE, xlab="OverallQual",
     main="Overall Quality")
curve(dgamma(x, shape = fd$estimate['shape'], rate = fd$estimate['rate']), 
      col="blue", add=TRUE)

```

Percentile
exponential pdf - 5th and 95th percentiles
```{r}
epb<- fitdistr(train$OverallQual, "exponential")
lambda <- epb$estimate
lambda
qexp(0.05,rate= lambda)
qexp(0.95, rate= lambda)
```
Confidential interval 
```{r}
CI(na.exclude(train$OverallQual), ci= 0.95)
```

### Modeling
```{r}
model = train[, which(sapply(train, function(x) sum(is.na(x))) == 0)]
ft = lm(train$SalePrice ~ train$OverallQual + train$YearBuilt + train$LotFrontage + train$GarageArea, data = train)
summary(ft)
```